---
title: "Template Analysis Urban Water Managment"
author: "Erin Cain"
date: "2/1/2022"
output: 
  html_document:
  theme: flatly
    # rmarkdown::github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
library(tidyverse)
library(dplyr)
library(formattable)
```

## City of Napa
 
## What Reports are we looking at? 

```{r}
report_1 <- "Urban Water Managment Plan"
report_2 <- "Water Loss Report"
report_3 <- "Conservation Report"
report_4 <- "Electronic Annual Report"
```

## What Metric are we comparing between the reports?

```{r}
metric <- "Volume Water Supplied in Acre Feet"
```

## What data are we using?

Tables Names: 

* Report 1: [Urban Water Managment Plan](https://wuedata.water.ca.gov/uwmp_export_2020.asp)
* Report 2: [Water Loss Report](https://wuedata.water.ca.gov/awwa_export)
* Report 3: [Conservation Report](https://www.waterboards.ca.gov/water_issues/programs/conservation_portal/conservation_reporting.html)
* Report 4: [Electronic Annual Report](https://www.waterboards.ca.gov/drinking_water/certlic/drinkingwater/eardata.html)

What agency are we focused on? 

```{r}
agency_of_interest <- "City of Napa"
```


Load in data: 
```{r, warning=FALSE, message=FALSE}
supply_and_demand_data <- read_rds("data/supply_and_demand_data.rds") %>% 
  filter(supplier_name %in% c("Napa  City Of", "Napa  City of", "NAPA, CITY OF"))
  
```

### How is our chosen metric described in the report data?

```{r}
volume_supplied_report_1 <- "VOLUME_OF_WATER_SUPPLIED_AF"
volume_supplied_report_2 <- "WS_WATER_SUPPLIED_VOL_AF"
volume_supplied_report_3 <- "REPORTED FINAL Total Potable Water Production"
volume_supplied_report_4 <- "WPAnnualTotal"
```


### Metrics {.tabset}

#### Urban Water Managment Plan

```{r}
uwmp_supply_table <- supply_and_demand_data %>% filter(report_name == "UWMP", category == "supply")
uwmp_total_supply <- sum(supply_and_demand_data %>% filter(report_name == "UWMP", category == "supply") %>% pull(volume_af), na.rm = T) # Summing together categories describes above
uwmp_total_supply

uwmp_metric <- uwmp_total_supply
```

#### Water Loss Report 

Water supplied subtracts exported water from own sources

```{r}
wlr_supply_table <- supply_and_demand_data %>% filter(report_name == "WLR", year == 2020, category == "supply" | category == "supply total")
wlr_total_supply <- sum(supply_and_demand_data %>% filter(report_name == "WLR", category == "supply", use_type != "ws_exported_vol_af", year == 2020) %>% pull(volume_af), na.rm = T) # Summing together categories describes above
wlr_total_supply_less_exported <- wlr_supply_table %>%
  filter(use_type == "ws_water_supplied_vol_af")
wlr_metric <- wlr_total_supply_less_exported$volume_af
```

#### Conservation Report 

```{r}
cr_supply_table <- supply_and_demand_data %>% filter(report_name == "CR", year == 2020, category == "supply" | category == "supply total")
cr_total_supply <- sum(cr_supply_table %>% pull(volume_af), na.rm = T)
cr_metric <- cr_total_supply
```

#### eAR

```{r}
ear_supply_table <- supply_and_demand_data %>% filter(report_name == "EAR", year == 2020, category == "supply total", !grepl("max", use_type))
ear_total_supply <- filter(ear_supply_table, use_type != "annualtotal") %>%
  mutate(volume_af = ifelse(use_type ==  "annualsold", volume_af*-1, volume_af)) %>%
  summarize(volume_af = sum(volume_af))
ear_metric <- ear_total_supply$volume_af
```

### Summarizing Metric Across Reports 

```{r}
supply_by_report <- tibble(
  "Agency" = rep("City of Napa", 4),
  "Report" = c("Urban Water Management Plan", "Water Loss Report", "Conservation Report", "Electronic Annual Report"),
  "Volume AF" = as.numeric(c(uwmp_metric,wlr_metric,cr_metric,ear_metric)),
  "Metric" = rep("Annual AF Supply", 4))
supply_by_report
write_rds(supply_by_report, "data/napa_supply_af_volumes.rds")
```

```{r}
library(ggplot2)
ggplot(supply_by_report, aes(y = Report, x = `Volume AF`, fill = Report)) +
  geom_col() + 
  labs(x = "Reported Annual AF Supply", y = "", 
       title = "Reported Annual Water Supply Across Reporting Requirements") +
  theme_minimal() +
  theme(legend.position="none", text = element_text(size=18)) + 
  scale_fill_manual(values = wesanderson::wes_palette("Royal2"))

```

### Delta of Metrics Across Reports {.tabset}

```{r}
total_water_supply_table <- 
  tibble("agency" = rep(c("City of Napa"), 6),
         "reports_compared" = c("UWMP & WLR", "UWMP & CR", "UWMP & EAR", 
                                "WLR & CR", "WLR & EAR", "CR & EAR"),
         "report_a" =  c("UWMP", "UWMP", "UWMP", "WLR", "WLR", "CR"),
         "report_b" =  c("WLR", "CR", "EAR", "CR", "EAR", "EAR"),
         "report_a_metric" = c(uwmp_metric, uwmp_metric, uwmp_metric,
                               wlr_metric, wlr_metric, cr_metric), 
         "report_b_metric" = c(wlr_metric, cr_metric, ear_metric, 
                               cr_metric, ear_metric, ear_metric),)

total_water_supply_af_deltas <- total_water_supply_table %>% 
  mutate(delta = abs(report_a_metric - report_b_metric), 
         percent_delta = abs((report_a_metric / report_b_metric - 1) * 100))

write_rds(total_water_supply_af_deltas, "data/napa_total_water_supply_af_deltas.rds")
```

### Delta Table 

The table below shows the delta between the reports

```{r, warning=FALSE}

clean_na = function(x){
  x %>%
    round(2) %>%
    as.character() %>%
    replace_na('-') %>%
    return()
}

delta_matrix <- total_water_supply_af_deltas %>% 
  select("Reports" = report_a, report_b, delta) %>%
  pivot_wider(names_from = "report_b", values_from = "delta")  %>%
  mutate_if(is.numeric, clean_na)


formattable(delta_matrix, list(area(col = 2:4) ~ color_tile("transparent", "pink")))
```

### % Delta Table 

The table below shows the percent delta between the reports

```{r, warning=FALSE}
percent_delta_matrix <- total_water_supply_af_deltas %>% 
  select("Reports" = report_a, report_b, percent_delta) %>%
  pivot_wider(names_from = "report_b", values_from = "percent_delta") %>%
  mutate_if(is.numeric, clean_na)

formattable(percent_delta_matrix, list(area(col = 2:4) ~ color_tile("transparent", "pink")))
```


## Investigating Differences 

```{r}
# read in lookup tables for use type categories
use_type_lookup <- read_rds("data/supply_type_lookup.rds")
```

The largest difference is between the eAR and the Water Loss Report.  

```{r}
all_reports_supply <- bind_rows(uwmp_supply_table, wlr_supply_table, cr_supply_table, ear_supply_table) %>% 
  left_join(use_type_lookup) %>%
  glimpse

write_rds(all_reports_supply, "data/napa_s_with_subcategories.rds")
```

#### Stacked Bar Plot: 
```{r}
# removed NA use groups from plot because the only one that is missing is "total"
filter(all_reports_supply, is.na(use_group))
ggplot(filter(all_reports_supply, !is.na(use_group), use_group != "total volume"), aes(y = report_name, x = volume_af, fill = use_group)) +
  geom_col() + 
  labs(x = "Reported Annual AF Supply", y = "", 
       title = "Reported Annual Water Supply Across Reporting Requirements") +
  theme_minimal() +
  theme(text = element_text(size=18)) 
```

