---
title: "Template Analysis Urban Water Managment"
author: "Erin Cain"
date: "2/1/2022"
output: 
  html_document:
  # rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
library(tidyverse)
library(dplyr)
library(formattable)
```

## City of Napa
 
## What Reports are we looking at? 

* "Urban Water Management Plan"
* "Water Loss Report"
* "Conservation Report"
* "Electronic Annual Report"

## What Metric are we comparing between the reports?

"Volume Water Demand in Acre Feet"

## What data are we using?

Tables Names: 

* Report 1: [Urban Water Managment Plan](https://wuedata.water.ca.gov/uwmp_export_2020.asp) (Table 4.1)
* Report 2: [Water Loss Report](https://wuedata.water.ca.gov/awwa_export)
* Report 3: [Conservation Report](https://www.waterboards.ca.gov/water_issues/programs/conservation_portal/conservation_reporting.html)
* Report 4: [Electronic Annual Report](https://www.waterboards.ca.gov/drinking_water/certlic/drinkingwater/eardata.html)

What agency are we focused on? 

"Napa City Of"

Load in data: 
```{r, warning=FALSE, message=FALSE}
supply_and_demand_data <- read_rds("../data/supply_and_demand_data.rds") %>% 
  filter(supplier_name %in% c("Napa  City Of", "Napa  City of", "NAPA, CITY OF"))
  
```

### How is our chosen metric described in the report data?

```{r}
volume_demand_report_1 <- "WATER_DEMAND_VOLUME_2020" # check that there is only one catagory, if not sum up
volume_demand_report_2 <- "AC_AUTH_CONSUMPTION_VOL_AF"
volume_demand_report_3 <- c("REPORTED FINAL Total Potable Water Production", "FINAL Percent Residential Use", 
                            "REPORTED FINAL Commercial Agricultural Water", "REPORTED FINAL Commercial, Industrial, and Institutional Water", 
                            "REPORTED Recycled Water", "REPORTED Non-Revenue Water") # Need to sum up reported metrics to get total demand 
volume_demand_report_4 <- "WDAnnualTotal"
```


### Metrics {.tabset}

#### Urban Water Managment Plan

UWMP Table 4.1 splits data into 9 categories:

* Multi-Family
* Commercial
* Institutional/Governmental
* Landscape
* Single Family
* Sales/Transfers/Exchanges to other agencies
* Other 
* Losses
* Agricultural Irrigation

```{r}
uwmp_demand_table <- supply_and_demand_data %>% filter(report_name == "UWMP", category == "demand" | category == "demand total" & 
                                                        use_type == "recycled water demand")
uwmp_total_demand <- sum(supply_and_demand_data %>% filter(report_name == "UWMP", category == "demand") %>% pull(volume_af), na.rm = T) # Summing together categories describes above
uwmp_total_demand
```

It DOES NOT include recycled water
```{r}
uwmp_recycled_water <- sum(supply_and_demand_data %>% filter(report_name == "UWMP", 
                                                        category == "demand total", 
                                                        use_type == "recycled water demand") %>% 
                        pull(volume_af), na.rm = T)
uwmp_recycled_water
```

```{r}
uwmp_metric <- uwmp_total_demand + uwmp_recycled_water
uwmp_metric
```


#### Water Loss Report 

The Water Loss Report only has one demand category: report 2 authorized consumption

```{r}
wlr_demand_table <- supply_and_demand_data %>% filter(report_name == "WLR", year == 2020, category == "demand" | 
                                                        category == "losses" & use_type == "wl_water_losses_vol_af" |
                                                        category == "supply"& use_type == "ws_exported_vol_af")
wlr_authorized_consumption <- supply_and_demand_data %>% filter(report_name == "WLR", year == 2020, category == "demand total") %>% pull(volume_af)
wlr_authorized_consumption
```

Water losses are not included in demand for the Water Loss Report. Losses are reported Separately and the value is below: 
```{r}
wlr_losses <- supply_and_demand_data %>% filter(report_name == "WLR", year == 2020, 
                                                category == "losses", use_type == "wl_water_losses_vol_af") %>% pull(volume_af)
wlr_losses
```

Water exported in AF are not included in demand for the Water Loss Report. Exports are reported separately and the value is below: 
```{r}
wlr_exports <- supply_and_demand_data %>% filter(report_name == "WLR", year == 2020, 
                                                 category == "supply", use_type == "ws_exported_vol_af") %>% pull(volume_af)
wlr_exports
```
```{r}
wlr_metric <- wlr_authorized_consumption + wlr_losses + wlr_exports
wlr_metric
```


#### Conservation Report 
The Conservation Report splits data into 5 categories: We summed together Residential Demand + Commercial Agriculture Demand + Commercial Industrial Institutional Demand, Recycled Demand, and Non Revenue Demand
```{r}
cr_supply_2020 <- supply_and_demand_data %>% 
  filter(report_name == "CR", year == 2020, category == "supply total") %>% 
  select(-use_type, -category) %>%
  rename("supply_total" = volume_af)

cr_demand_table <- supply_and_demand_data %>% 
  filter(report_name == "CR", year == 2020, category == "demand") %>% 
  left_join(cr_supply_2020) %>%
  mutate(volume_af = ifelse(use_type == "final percent residential use", (volume_af / 100) * supply_total, volume_af),
         use_type = ifelse(use_type == "final percent residential use", "final residential use", use_type)) %>%
  select(-supply_total) %>% 
  group_by(use_type) %>%
  summarise(volume_af = sum(volume_af, na.rm = T)) %>%
  ungroup() %>%
  mutate(report_name = "CR", 
         supplier_name = "Napa City of",
         year = 2020, 
         category = "demand")
  

residential_demand <- sum(supply_and_demand_data %>% 
                            filter(report_name == "CR", year == 2020, 
                                 category == "supply total") %>% 
                            pull(volume_af) * 
                            (supply_and_demand_data %>% 
                            filter(report_name == "CR", year == 2020, 
                                 category == "demand", use_type == "final percent residential use") %>% 
                            pull(volume_af))/100, na.rm = T) 

commercial_agriculture_demand <- sum(supply_and_demand_data %>% filter(report_name == "CR", year == 2020, 
                                                                   category == "demand", 
                                                                   use_type == "reported final commercial agricultural water") %>% pull(volume_af), na.rm = T)

commercial_industrial_institutional_demand <- sum(supply_and_demand_data %>% filter(report_name == "CR", year == 2020, 
                                                                   category == "demand", 
                                                                   use_type == "reported final commercial, industrial, and institutional water") %>% pull(volume_af), na.rm = T)

recycled_demand <- sum(supply_and_demand_data %>% filter(report_name == "CR", year == 2020, 
                                                                   category == "demand", 
                                                                   use_type == "reported recycled water") %>% pull(volume_af), na.rm = T)

non_revenue_demand <- sum(supply_and_demand_data %>% filter(report_name == "CR", year == 2020, 
                                                                   category == "demand", 
                                                                   use_type == "reported non-revenue water") %>% pull(volume_af), na.rm = T)
# Units are in AF
cr_metric <-  sum(residential_demand, commercial_agriculture_demand, commercial_industrial_institutional_demand, recycled_demand, non_revenue_demand, na.rm = T)
cr_metric 
```

#### eAR
The EAR reports total Water Demand but also has it broken down by the following categories: 

* Multi-Family 
* Single Family
* Commercial Industrial 
* Industrial
* Landscape Irrigation 
* Agriculture 
* Percent Recycled 
* Other


```{r}
ear_demand_table <- supply_and_demand_data %>% 
  filter(report_name == "EAR", year == 2020, category == "demand", use_type != "Total") %>% 
  group_by(use_type) %>%
  summarise(volume_af = sum(volume_af, na.rm = T)) %>%
  ungroup() %>%
  mutate(report_name = "EAR", 
         supplier_name = "NAPA, CITY OF",
         year = 2020, 
         category = "demand")

ear_metric <-  supply_and_demand_data %>% filter(report_name == "EAR", year == 2020, 
                                                 category == "demand total", 
                                                 use_type == "annualtotal") %>% pull(volume_af)

ear_metric
```

### Summarizing Metric Across Reports 

```{r}
demand_by_report <- tibble(
  "Agency" = rep("City of Napa", 4),
  "Report" = c("Urban Water Management Plan", "Water Loss Report", "Conservation Report", "Electronic Annual Report"),
  "Volume AF" = as.numeric(c(uwmp_metric,wlr_metric,cr_metric,ear_metric)),
  "Metric" = rep("Annual Total Demand", 4))
demand_by_report
write_rds(demand_by_report, "../data/napa_demand_af_volumes.rds")
```

```{r}
library(ggplot2)
ggplot(demand_by_report, aes(y = Report, x = `Volume AF`, fill = Report)) +
  geom_col() + 
  labs(x = "Reported Annual AF Demand", y = "", 
       title = "Reported Annual Water Demand Across Reporting Requirements") +
  theme_minimal() +
  theme(legend.position="none", text = element_text(size=18)) + 
  scale_fill_manual(values = wesanderson::wes_palette("Royal2"))

```

### Delta of Metrics Across Reports {.tabset}

```{r}
total_water_demand_table <- 
  tibble("agency" = rep(c("City of Napa"), 6),
         "reports_compared" = c("UWMP & WLR", "UWMP & CR", "UWMP & EAR", 
                                "WLR & CR", "WLR & EAR", "CR & EAR"),
         "report_a" =  c("UWMP", "UWMP", "UWMP", "WLR", "WLR", "CR"),
         "report_b" =  c("WLR", "CR", "EAR", "CR", "EAR", "EAR"),
         "report_a_metric" = c(uwmp_metric, uwmp_metric, uwmp_metric,
                               wlr_metric, wlr_metric, cr_metric), 
         "report_b_metric" = c(wlr_metric, cr_metric, ear_metric, 
                               cr_metric, ear_metric, ear_metric),)

total_water_demand_af_deltas <- total_water_demand_table %>% 
  mutate(delta = abs(report_a_metric - report_b_metric), 
         percent_delta = abs((report_a_metric / report_b_metric - 1) * 100))

write_rds(total_water_demand_af_deltas, "../data/napa_total_water_demand_af_deltas.rds")
```



### Delta Table 

The table below shows the delta between the reports

```{r, warning=FALSE}

clean_na = function(x){
  x %>%
    round(2) %>%
    as.character() %>%
    replace_na('-') %>%
    return()
}

delta_matrix <- total_water_demand_af_deltas %>% 
  select("Reports" = report_a, report_b, delta) %>%
  pivot_wider(names_from = "report_b", values_from = "delta")  %>%
  mutate_if(is.numeric, clean_na)


formattable(delta_matrix, list(area(col = 2:4) ~ color_tile("transparent", "pink")))
```

### % Delta Table 

The table below shows the percent delta between the reports

```{r, warning=FALSE}
percent_delta_matrix <- total_water_demand_af_deltas %>% 
  select("Reports" = report_a, report_b, percent_delta) %>%
  pivot_wider(names_from = "report_b", values_from = "percent_delta") %>%
  mutate_if(is.numeric, clean_na)

formattable(percent_delta_matrix, list(area(col = 2:4) ~ color_tile("transparent", "pink")))
```


## Investigating Differences 

```{r}
# read in lookup tables for use type categories
use_type_lookup <- read_rds("../data/use_type_lookup.rds")
```

The largest difference is between the CR and the Water Loss Report. 

```{r}
all_reports_demand <- bind_rows(uwmp_demand_table, wlr_demand_table, cr_demand_table, ear_demand_table) %>% 
  left_join(use_type_lookup) %>%
  mutate(supplier_name = "City of Napa") %>%
  glimpse

napa_d_with_subcategories <- write_rds(all_reports_demand, "../data/napa_d_with_subcategories.rds")
```

#### Stacked Bar Plot: 
```{r}
# removed NA use groups from plot because the only one that is missing is "total"
filter(all_reports_demand, is.na(use_group))
ggplot(filter(all_reports_demand, !is.na(use_group)), aes(y = report_name, x = volume_af, fill = use_group)) +
  geom_col() + 
  labs(x = "Reported Annual AF Demand", y = "", 
       title = "Reported Annual Water Demand Across Reporting Requirements") +
  theme_minimal() +
  theme(text = element_text(size=18)) 
```

Discussion of Observations: 
- Refer to definitions 

*Sales transfer exchanges* 
- Appear to be similar for WLR and EAR; UWMP is smaller, why? None for CR

*Residential*
- Very similar for UWMP, EAR, and CR

*Recycled*
- UWMP has the highest; little to none for others

*Losses*
- UWMP and WLR are similar though WLR is larger

*Landscape*
- UWMP and EAR are very similar

*Commercial industrial institutional*
- UWMP and EAR are very similar; CR is a little smaller, why?

*Agricultural irrigation*
- UWMP and EAR are very similar

**Insights**
- One reason why EAR may be lower than WLR and UWMP is that losses are not reported
in the EAR
- CR may be lower due to data quality issues - reported 0 for non-revenue water
(aka losses), 0 for recycled water, and 0 for agriculture irrigation.