---
title: "Template Analysis Urban Water Managment"
author: "Erin Cain"
date: "2/1/2022"
output: 
  html_document:
  theme: flatly
    # rmarkdown::github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=12, fig.height=8)
library(tidyverse)
library(dplyr)
library(formattable)
```

## City of Napa
 
## What Reports are we looking at? 

```{r}
report_1 <- "Urban Water Managment Plan"
report_2 <- "Water Loss Report"
report_3 <- "Conservation Report"
report_4 <- "Electronic Annual Report"
```

## What Metric are we comparing between the reports?

```{r}
metric <- "Volume Water Supplied in Acre Feet"
```

## What data are we using?

Tables Names: 

* Report 1: [Urban Water Managment Plan](https://wuedata.water.ca.gov/uwmp_export_2020.asp)
* Report 2: [Water Loss Report](https://wuedata.water.ca.gov/awwa_export)
* Report 3: [Conservation Report](https://www.waterboards.ca.gov/water_issues/programs/conservation_portal/conservation_reporting.html)
* Report 4: [Electronic Annual Report](https://www.waterboards.ca.gov/drinking_water/certlic/drinkingwater/eardata.html)

What agency are we focused on? 

```{r}
agency_of_interest <- "Sante Fe Irrigarion District"
```


Load in data: 
```{r, warning=FALSE, message=FALSE}
data_report_1 <- readxl::read_excel("../data-raw/uwmp_table_2_2_r_conv_to_af.xlsx") %>% 
  filter(WATER_SUPPLIER_NAME == "Napa  City Of")
  
data_report_2 <- readxl::read_excel("../data-raw/water_audit_data_conv_to_af.xlsx") %>% 
  filter(REPORTING_YEAR == 2020, WATER_SUPPLIER_NAME == "Napa  City Of") 


data_report_3 <- readxl::read_excel("../data-raw/conservation-report-uw-supplier-data120721.xlsx") %>% 
  mutate(year = lubridate::year(`Reporting Month`)) %>%
  filter(year == 2020, `Supplier Name` == "Napa  City of")


data_report_4 <-  read.delim("../data-raw/EAR_ResultSet_2020RY.txt")

data_report_4 <- data_report_4 %>%
  filter(WSSurveyID == 428709) %>%
  filter(SurveyName == "2020 EAR", SectionID %in% c("06 Supply-Delivery", "01 Intro")) 

```

### How is our chosen metric described in the report data?

```{r}
volume_supplied_report_1 <- "VOLUME_OF_WATER_SUPPLIED_AF"
volume_supplied_report_2 <- "WS_WATER_SUPPLIED_VOL_AF"
volume_supplied_report_3 <- "REPORTED FINAL Total Potable Water Production"
volume_supplied_report_4 <- "WPAnnualTotal"
```


### Metrics {.tabset}

#### Urban Water Managment Plan

```{r}
metric_report_1 <- data_report_1 %>% pull(volume_supplied_report_1)
metric_report_1
```

#### Water Loss Report 

```{r}
metric_report_2 <-  data_report_2 %>% pull(volume_supplied_report_2)
metric_report_2 

```

#### Conservation Report 

```{r}
#Check Units
data_report_3 %>% pull(`Water Production Units`) 

# Units are in AF
metric_report_3 <-  sum(data_report_3 %>% pull(volume_supplied_report_3))
metric_report_3 
```

#### eAR

```{r}
# Check Units 
data_report_4 %>% filter(QuestionName == "WPUnitsofMeasure") %>% pull(QuestionResults)


scale_to_AF <- 3.06888
# Units are in MG, multiply by scaling factor to convert to AF
metric_report_4 <-  as.numeric(data_report_4 %>% 
                                 filter(QuestionName == "WPAnnualTotal") %>%
                                 pull(QuestionResults)) * scale_to_AF
metric_report_4 
```

### Summarizing Metric Across Reports 

```{r}
supply_by_report <- tibble("Report" = c(report_1, report_2, report_3, report_4),
                           "Annual AF Supply" = as.numeric(c(metric_report_1, 
                                                             metric_report_2, 
                                                             metric_report_3, 
                                                             metric_report_4)))
```

```{r}
library(ggplot2)
ggplot(supply_by_report, aes(y = Report, x = `Annual AF Supply`, fill = Report)) +
  geom_col() + 
  labs(x = "Reported Annual AF Supply", y = "", 
       title = "Reported Annual Water Supply Across Reporting Requirements") +
  theme_minimal() +
  theme(legend.position="none", text = element_text(size=18)) + 
  scale_fill_manual(values = wesanderson::wes_palette("Royal2"))

```

### Delta of Metrics Across Reports {.tabset}

```{r}
total_water_supply_table <- 
  tibble("agency" = rep(c("City of Napa"), 6),
         "reports_compared" = c("UWMP & WLR", "UWMP & CR", "UWMP & EAR", 
                                "WLR & CR", "WLR & EAR", "CR & EAR"),
         "report_a" =  c("UWMP", "UWMP", "UWMP", "WLR", "WLR", "CR"),
         "report_b" =  c("WLR", "CR", "EAR", "CR", "EAR", "EAR"),
         "report_a_metric" = c(metric_report_1, metric_report_1, metric_report_1, 
                               metric_report_2, metric_report_2, metric_report_3), 
         "report_b_metric" = c(metric_report_2, metric_report_3, metric_report_4, 
                               metric_report_3, metric_report_4, metric_report_4),)

total_water_supply_af_deltas <- total_water_supply_table %>% 
  mutate(delta = abs(report_a_metric - report_b_metric), 
         percent_delta = abs((report_a_metric / report_b_metric - 1) * 100))

write_rds(total_water_supply_af_deltas, "../data/napa_total_water_supply_af_deltas.rds")
```

### Delta Table 

The table below shows the delta between the reports

```{r, warning=FALSE}

clean_na = function(x){
  x %>%
    round(2) %>%
    as.character() %>%
    replace_na('-') %>%
    return()
}

delta_matrix <- total_water_supply_af_deltas %>% 
  select("Reports" = report_a, report_b, delta) %>%
  pivot_wider(names_from = "report_b", values_from = "delta")  %>%
  mutate_if(is.numeric, clean_na)


formattable(delta_matrix, list(area(col = 2:4) ~ color_tile("transparent", "pink")))
```

### % Delta Table 

The table below shows the percent delta between the reports

```{r, warning=FALSE}
percent_delta_matrix <- total_water_supply_af_deltas %>% 
  select("Reports" = report_a, report_b, percent_delta) %>%
  pivot_wider(names_from = "report_b", values_from = "percent_delta") %>%
  mutate_if(is.numeric, clean_na)

formattable(percent_delta_matrix, list(area(col = 2:4) ~ color_tile("transparent", "pink")))
```


## Investigating Differences 

The largest difference is between the eAR and the Water Loss Report. 

TODO Why...?

