---
title: "Guided examples & key insights"
output:
  rmarkdown::html_document:
    theme: yeti
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width=12, fig.height=8)
library(tidyverse)
library(knitr)
library(DT)
```

```{r data_pull, include = F}
# City of Napa, Santa Fe Irrigation District, Moulton Niguel
selected_ids <- c("CA2810003", "CA3710023", "CA3010073")
supply_and_demand_data <- read_rds("../data/supply_and_demand_data.rds") %>% 
  filter(supplier_id %in% selected_ids)
data_dictionary <- read_csv("../data/data_dictionary.csv")
# decided to include sold (EAR) and exported water (WLR) to demand use types
demand_data_raw <- filter(supply_and_demand_data, grepl("demand", category) | category == "losses" | use_type %in% c("sold", "ws_exported_vol_af"))
supply_data_raw <- filter(supply_and_demand_data, grepl("supply", category) | !use_type %in% c("sold", "ws_exported_vol_af"))

supply_total <- supply_and_demand_data %>% 
  filter(category == "supply total") %>% 
  select(-use_type, -category) %>%
  rename("supply_total" = volume_af)

# residential use for conservation report is reported as a percentage
demand_data <- demand_data_raw %>% 
  left_join(supply_total) %>%
  mutate(volume_af = ifelse(use_type == "final percent residential use", (volume_af / 100) * supply_total, volume_af),
         use_type = ifelse(use_type == "final percent residential use", "final residential use", use_type)) %>%
  select(-supply_total) %>% 
  # sum across all months
  group_by(report_name, supplier_id, year, category, use_type) %>%
  summarise(volume_af = sum(volume_af, na.rm = T))

supply_data <- supply_data_raw %>%
  # sum across all months
  group_by(report_name, supplier_id, year, category, use_type) %>%
  summarise(volume_af = sum(volume_af, na.rm = T))

# use_type "op" and "sold" should be the same so remove op
demand_use_type <- filter(demand_data, category != "demand total", use_type != "total", use_type != "op") %>%
  mutate(category = "demand")
supply_use_type <- filter(supply_data, (category == "supply" | use_type == "reported final total potable water production"), use_type != "total")

# Lookup tables ####
# use type lookup
use_type_lookup <- read_rds("../data/use_type_lookup.rds") %>%
  mutate(category = "demand")

# supply type lookup
supply_type_lookup <- read_rds("../data/supply_type_lookup.rds") %>%
  mutate(category = "supply")

filter(demand_data, category == "demand") %>% ungroup() %>% distinct(use_type)
filter(supply_data, category == "supply") %>% ungroup() %>% distinct(use_type)
```

As part of the Urban Water Data Reporting pilot project, the California Water 
Data Consortium (Consortium) will identify opportunities to align current water 
supply and use data collected and reported by local and wholesale water agencies 
with State agency reporting requirements to improve data use, reduce reporting burden, 
and expand access to timely, interoperable, high-quality, secure, reliable data. 
In addition to improving public understanding and access to core water supply and 
use data, these data services will provide the source data needed to generate 
critical information for managing water resources.

FlowWest is working with the Consortium to identify opportunities for alignment 
through quantitative analysis, interviews, and a workshop. This document summarizes 
the key insights gathered from quantitative analysis of urban water use and supply 
data reported on the Urban Water Management Plan (UWMP), Monthly Urban Water 
Conservation Report (CR), Electronic Annual Report (EAR), and Water Loss Audit (WLR), 
and through interviews with local retail and wholesale agencies. This analysis 
was not designed to be complete and inclusive of all aspects of water use and 
supply data reported. Rather, it was designed to guide additional conversations 
and analysis surrounding state urban water use and supply data.

# Data mapping {.tabset}

State reporting requirements include definitions for the water use and supply
groups that suppliers are required to report on. In many cases, the name of the 
water use or supply group varies across reporting requirements and the definition
may also vary.

The creation of an inclusive list of water use and supply groups, and mappings across
reporting requirements was informed by the synthesis of summary documents for 
state reporting requirements (also referred to as templates). Refer to the Data Mapping
tab for additional information.

```{r demand_data, include = F, echo = F}
demand_analysis_raw <- demand_use_type %>%
  left_join(use_type_lookup) %>%
  filter(year == 2020) %>%
  mutate(year = as.factor(year),
         supplier_id = case_when(supplier_id == "CA2810003" ~ "Agency 1",
                                 supplier_id == "CA3710023" ~ "Agency 2",
                                 supplier_id == "CA3010073" ~ "Agency 3")) 
demand_analysis <- demand_analysis_raw %>%
  group_by(supplier_id, report_name, use_group, year) %>%
  summarize(volume_af = sum(volume_af))

total_demand <- filter(demand_analysis_raw, use_type != "recycled", !is.na(use_group)) %>%
  group_by(supplier_id, report_name) %>%
  summarize(total_demand = sum(volume_af))

demand_analysis_percent <- filter(demand_analysis_raw, use_type != "recycled", !is.na(use_group)) %>%
  group_by(supplier_id, report_name, use_group, year) %>%
  summarize(volume_af = sum(volume_af)) %>%
  left_join(total_demand) %>%
  mutate(percent_total = volume_af/total_demand)
  
# 10 demand use groups
use_group_list <- unique(demand_analysis$use_group)

# report_colors <- c("WLR" = "#1b9e77",
#                    "CR" = "#d95f02",
#                    "EAR" = "#7570b3",
#                    "UWMP" = "#e7298a")

report_colors <- c("WLR" = "#9A8822",
                   "CR" = "#74A089",
                   "EAR" = "#F8AFA8",
                   "UWMP" = "#FDDDA0")

use_type_stacked_bar <- function(dat) {
  ggplot(dat, aes(x = year, y = volume_af, fill = report_name)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = report_colors) +
  facet_wrap(~supplier_id, scales = "free_y") +
  labs(x = "") +
  theme_minimal() +
  theme(legend.title = element_blank())
}

use_type_percent_bar <- function(dat) {
  ggplot(dat, aes(x = year, y = percent_total, fill = report_name)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = report_colors) +
  facet_wrap(~supplier_id, scales = "free_y") +
  labs(x = "") +
  theme_minimal() +
  theme(legend.title = element_blank())
}
```

# Analysis themes {.tabset}

## Disconnect between data available and reporting requirements

The CR requires suppliers to report water use and supply data monthly and the EAR
requires an annual submission of data aggregated by month. The UWMP is required 
every five years and the WLR is required annually.

Monthly reporting of water consumption data, particularly for the CR, is challenging 
because some suppliers may collect consumption data based on bimonthly meter read
cycles or may not collect data for a particular category. The quotes from two retailers
below describe the challenges in meeting monthly reporting requirements for water use data
and the different approaches that are used to meet the requirements.

*"At the customer level, it's every two months; bimonthly for most customers. So you don't
have real time data for that; and so, there's a little bit of an issue when, like we
do a monthly report and they want to know residential use. We use, you know, an equation
to guess based on a ratio from last year or something to determine that; we just don't have
real time data for metered billed use. But we have it for production, so there's a little bit
of inconsistency there in terms of the time, the time periods that are available for us."* 
(Source: Large retailer 2022).

*"What they are reporting to the state is not what is being asked for. And it is pretty common
knowledge among water agencies that's not what you're doing. We don't report water use at all.
Honestly, we report production, and we are hoping to change that, but we honestly just don't
have the staff or the capabilities to do it accurately."*
(Source: Small retailer 2022).

In Figure 1, percent of total water use reported by water use category is compared across
the CR, UWMP, EAR, and WLR. The CR asks for water use to be reported by each of the
categories shown (residential, agricultural and irrigation, CII, and losses) though 
that data is not always reported even when it exists for other reports. For instance,
Agency 1 reports agricultural use and losses on the EAR and UWMP but reports neither 
on the CR. While differences exist across all reports, the CR is often the outlier, 
and where the biggest differences exist.

```{r disconnect, echo = F, fig.cap = "Figure 1. Percent of total water use reported by water use category (only those reported on the CR) compared across the CR, UWMP, EAR, and WLR for three agencies."}
filter(demand_analysis_percent, use_group %in% c("residential", "agricultural irrigation",
                                         "commerical industrial institutional", "losses")) %>%
  mutate(use_group = case_when(use_group == "residential" ~ "Residential",
                               use_group == "agricultural irrigation" ~ "Agricultural",
                               use_group == "commerical industrial institutional" ~ "CII",
                               use_group == "losses" ~ "Losses"),
         percent_total = percent_total * 100) %>%
  ggplot(aes(x = year, y = percent_total, fill = report_name)) +
  geom_bar(stat = "identity", position = position_dodge(width = 1, preserve = "single")) +
  scale_fill_manual(values = report_colors) +
  facet_grid(use_group~supplier_id, scales = "free_y") +
  labs(x = "",
       y = "Percent of total use reported") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        strip.text.y = element_text(size = 8)) 

```

This suggests that for some suppliers data reported on the CR represents an estimate
rather than actual volumes. This characteristic distinguishes the CR from the UWMP,
EAR, and WLR. 

The disconnect between state reporting requirements and the reality of local supplier
data collection and analysis methods extends beyond the CR. For instance, one supplier
discussed the timing expectations of the upcoming Annual Water Supply and Demand Assessment:

*"Now have the Annual Water Supply and Demand Assessment Reporting requirement, right.
This was happening July 1st and and they assume that every district does it the same 
way, you know, we're all gonna know our projection end of June or all this for the next 
year and stuff. And we don't do it that way."*
(Source: Larger retailer 2022).

## Redundancy across reports

The tables presented in the Data Mapping tab provide an interpretation of the alignment in data elements across
reports. While the names and definitions may vary, the data reported are very similar
indicating that for all intents and purposes they are interpreted as the same. 
Regardless of the intention of reporting requirements, this suggests that there 
is redundancy across reports. 

This redundancy may not always be apparent when comparing raw volumes because there may
be differences in the exact values being reported on each report as discussed above.
When comparing use types across reports by percentage of total demand, in some cases,
the redundancy becomes more apparent.

For instance, Agency 1 reports different volumes of commercial, industrial and institutional
consumption across reports where the values reported on the EAR and UWMP are more similar than
on the CR (Figure 2). CII demand reported on the CR is 23% different than the UWMP and 20% different than the EAR.
However, when comparing this use type as a percentage of total demand reported, all 
reports are very similar, within 1% difference (Figure 2).

```{r redundancy, echo = F, fig.cap = "Figure 2. Commercial, industrial, institutional (CII) volume in acre feet and as a percentage of total demand for Agency 1 on the CR, EAR, and UWMP."}
example_1_raw <- filter(demand_analysis, use_group == "commerical industrial institutional", supplier_id == "Agency 1") %>%
  ggplot(aes(x = year, y = volume_af, fill = report_name)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = report_colors) +
  labs(y = "Commerical industrial institutional volume (af)") +
  labs(x = "") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "none")

example_1_percent <- filter(demand_analysis_percent, use_group == "commerical industrial institutional", supplier_id == "Agency 1") %>%
  ggplot(aes(x = year, y = percent_total, fill = report_name)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = report_colors) +
  labs(y = "Percent of Total Demand: Commerical industrial institutional volume (af) ") +
  labs(x = "") +
  theme_minimal() +
  theme(legend.title = element_blank())

gridExtra::grid.arrange(example_1_raw, example_1_percent, nrow = 1)

```

```{r, include = F}
example_1 <- filter(demand_analysis_percent, use_group == "residential", supplier_id == "Agency 2") 

example_1_format <- select(example_1, report_name, volume_af) %>%
  pivot_wider(names_from = "report_name", values_from = "volume_af") %>%
  mutate(type = "volume_af") %>%
  bind_rows(select(example_1, report_name, percent_total) %>%
  pivot_wider(names_from = "report_name", values_from = "percent_total") %>%
  mutate(type = "percent_total")) %>%
  mutate(percent_diff = (abs(CR-UWMP)/(0.5 * (CR+UWMP)))*100)

example_2 <- filter(demand_analysis_percent, use_group == "commerical industrial institutional", supplier_id == "Agency 1")

example_2_format <- select(example_2, report_name, volume_af) %>%
  pivot_wider(names_from = "report_name", values_from = "volume_af") %>%
  mutate(type = "volume_af") %>%
  bind_rows(select(example_2, report_name, percent_total) %>%
  pivot_wider(names_from = "report_name", values_from = "percent_total") %>%
  mutate(type = "percent_total")) %>%
  mutate(percent_diff_cr_uwmp = (abs(CR-UWMP)/(0.5 * (CR+UWMP)))*100,
         percent_diff_cr_ear = (abs(CR-EAR)/(0.5 * (CR+EAR)))*100,
         percent_diff_ear_uwmp = (abs(EAR-UWMP)/(0.5 * (EAR+UWMP)))*100)
```
