---
title: "Report Combinations and Metric Options"
author: "Erin Cain"
date: "2/1/2022"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reports 

There are 5 different reports that must be compared with each other. There are 10 total comparisons that must be done for each metric. Each comparison is shown below: 


| UWMP| Conservation| eAR| Water Loss| Supply & Demand| Water Use Objective| 
| -------------- | -------- | ------- | ------------ | ----------------| ---------- | 
| **Conservation** | - | 1 | 2 | 3 | 4 | 
| **eAR** | - | - | 5 | 6 | 7 | 
| **Water Loss** | - | - | - | 8 | 9 | 
| **Supply & Demand ?** | - | - | - | - | 10 | 
|**Water Use Objective ?** | - | - | - | - | - | 

Can we do Supply & Demand or Water Use Objective. I think we will have to hold off on these until there is data
## Metrics

```{r}
# Conservation Report Metrics
readr::read_rds("data/conservation_report_fields.rds")

# Some of UWMP Metrics (many more tables we could pull from)
readr::read_rds("data/UWMP_fields.rds")

# Water Loss Report Metrics
readr::read_rds("data/WLR_fields.rds")

```
