---
title: "Template Analysis Urban Water Managment"
author: "Erin Cain"
date: "2/1/2022"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)

```

## Template Analysis for Urban Water Managment 

Currently we can use this template to streamline analysis BUT there is still a lot of manual work. It would be ideal if we could map through all the reports/metrics/agencies efficiently BUT before we can do that we need to have all metric pairs and report pairs generated. 

TODO discuss which approach makes most sense given quantity of analyzes we hope to get through. 
 
## What Reports are we looking at? 

```{r}
report_1 <- "Urban Water Managment Plan"
report_2 <- "Water Loss Report"
```

## What Metric are we comparing between the reports?

```{r}
metric <- "Volume Water Supplied in Acre Feet"
```

## What data are we using?

Tables Names: 

* Report 1: (source)
* Report 2: (source)

What agency are we focused on? To pick this we looked across both reports and looked for an agency that participates in both. 

```{r}
agency_of_interest <- "Healdsburg  City Of"
```


Load in data: 
```{r, warning=FALSE, message=FALSE}
data_report_1 <- readxl::read_excel("../data-raw/uwmp_table_2_2_r_conv_to_af.xlsx") %>% 
  filter(WATER_SUPPLIER_NAME == "Healdsburg  City Of")
data_report_1
  
data_report_2 <- readxl::read_excel("../data-raw/water_audit_data_conv_to_af.xlsx") %>% 
  filter(REPORTING_YEAR == 2020, WATER_SUPPLIER_NAME == "Healdsburg  City Of")
data_report_2
```

### How is our chosen metric described in the report data?

```{r}
volume_supplied_report_1 <- "VOLUME_OF_WATER_SUPPLIED_AF"
volume_supplied_report_2 <- "WS_WATER_SUPPLIED_VOL_AF"
```


```{r}
metric_report_1 <- data_report_1 %>% pull(volume_supplied_report_1)
metric_report_2 <-  data_report_2 %>% pull(volume_supplied_report_2)

delta_water_supplied <- metric_report_1 - metric_report_2
delta_water_supplied_percent <- (metric_report_1 / metric_report_2 - 1) * 100
```

The *difference* in the `r metric` between the `r report_1` and `r report_2` is: `r delta_water_supplied`. 

The *percent difference* is: `r delta_water_supplied_percent` %.


