---
title: "Template Analysis Urban Water Reporting"
author: "Erin Cain"
date: "2/1/2022"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)

```

## Template Analysis for Urban Water Reporting 

Template for streamlining analysis of urban water data reported. Fill out each section of the template below to define the reports, metric, and agency to compare across. The purpose of this analysis is to find the difference and percent difference of metrics reported to multiple reports. 
 
## What Reports are we looking at? 

```{r}
report_1 <- "Urban Water Managment Plan"
report_2 <- "Water Loss Report"
```

## What Metric are we comparing between the reports?

```{r}
metric <- "Volume Water Supplied in Acre Feet"
```

## What data are we using?

Tables Names: 

* Report 1: (source)
* Report 2: (source)

What agency are we focused on? To pick this we looked across both reports and looked for an agency that participates in both. 

```{r}
agency_of_interest <- "Sante Fe Irrigarion District"
```


Load in data: 
```{r, warning=FALSE, message=FALSE}
data_report_1 <- readxl::read_excel("../data-raw/uwmp_table_2_2_r_conv_to_af.xlsx") %>% 
  filter(WATER_SUPPLIER_NAME == "Santa Fe Irrigation District")
data_report_1
  
data_report_2 <- readxl::read_excel("../data-raw/water_audit_data_conv_to_af.xlsx") %>% 
  filter(REPORTING_YEAR == 2020, WATER_SUPPLIER_NAME == "Santa Fe Irrigation District")
data_report_2
```

### How is our chosen metric described in the report data?

```{r}
volume_supplied_report_1 <- "VOLUME_OF_WATER_SUPPLIED_AF"
volume_supplied_report_2 <- "WS_WATER_SUPPLIED_VOL_AF"
```


```{r}
metric_report_1 <- data_report_1 %>% pull(volume_supplied_report_1)
metric_report_1
metric_report_2 <-  data_report_2 %>% pull(volume_supplied_report_2)
metric_report_2 

delta_water_supplied <- metric_report_1 - metric_report_2
delta_water_supplied
delta_water_supplied_percent <- (metric_report_1 / metric_report_2 - 1) * 100
```

The *difference* in the `r metric` between the `r report_1` and `r report_2` is: `r delta_water_supplied` Acre Feet. 

The *percent difference* is: `r delta_water_supplied_percent` %.

## If % differnce is significant (define what significant is), why is it different?

