---
title: "Use and supply by category"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(DT)
```

```{r data_pull, include = F}
# City of Napa, Santa Fe Irrigation District, Moulton Niguel
selected_ids <- c("CA2810003", "CA3710023", "CA3010073")
supply_and_demand_data <- read_rds("data/supply_and_demand_data.rds") %>% 
  filter(supplier_id %in% selected_ids)
data_dictionary <- read_csv("data/data_dictionary.csv")
# decided to include sold (EAR) and exported water (WLR) to demand use types
demand_data_raw <- filter(supply_and_demand_data, grepl("demand", category) | category == "losses" | use_type %in% c("sold", "ws_exported_vol_af"))
supply_data_raw <- filter(supply_and_demand_data, grepl("supply", category) | !use_type %in% c("sold", "ws_exported_vol_af"))

supply_total <- supply_and_demand_data %>% 
  filter(category == "supply total") %>% 
  select(-use_type, -category) %>%
  rename("supply_total" = volume_af)

# residential use for conservation report is reported as a percentage
demand_data <- demand_data_raw %>% 
  left_join(supply_total) %>%
  mutate(volume_af = ifelse(use_type == "final percent residential use", (volume_af / 100) * supply_total, volume_af),
         use_type = ifelse(use_type == "final percent residential use", "final residential use", use_type)) %>%
  select(-supply_total) %>% 
  # sum across all months
  group_by(report_name, supplier_id, year, category, use_type) %>%
  summarise(volume_af = sum(volume_af, na.rm = T))

supply_data <- supply_data_raw %>%
  # sum across all months
  group_by(report_name, supplier_id, year, category, use_type) %>%
  summarise(volume_af = sum(volume_af, na.rm = T))

# use_type "op" and "sold" should be the same so remove op
demand_use_type <- filter(demand_data, category != "demand total", use_type != "total", use_type != "op") %>%
  mutate(category = "demand")
supply_use_type <- filter(supply_data, (category == "supply" | use_type == "reported final total potable water production"), use_type != "total")

# Lookup tables ####
# use type lookup
use_type_lookup <- read_rds("data/use_type_lookup.rds") %>%
  mutate(category = "demand")

# supply type lookup
supply_type_lookup <- read_rds("data/supply_type_lookup.rds") %>%
  mutate(category = "supply")

filter(demand_data, category == "demand") %>% ungroup() %>% distinct(use_type)
filter(supply_data, category == "supply") %>% ungroup() %>% distinct(use_type)
```

# Data mapping {.tabset}

State reporting requirements include definitions for the water use and supply
groups that suppliers are required to report on. In many cases, the name of the 
water use or supply group varies across reporting requirements and the definition
may also vary.

The creation of an inclusive list of water use and supply groups, and mappings across
reporting requirements was informed by the synthesis of summary documents for 
state reporting requirements (also referred to as templates).

## Water use groups

Table 1 below shows the mapping for water use groups across the Urban Water Management Plan (UWMP),
Monthly Urban Water Conservation Report (CR), Electronic Annual Report (EAR), and Water Loss Audit (WLR). 
The Water Use Objective and Annual Water Supply and Demand Assessment are upcoming 
requirements and are not included in this analysis. 

The UWMP divides water use into the most categories. The CR and EAR require
some of the same groups as the UWMP but not all. The WLR has the least amount of
overlap, in terms of water use groups, across reports.

```{r use_mapping, echo = F}
# Assumptions are the mappings created for the categories
# multiple terms per use group so errors and creates list
#TODO make this look nicer
use_type_mapping_table <- filter(use_type_lookup, !is.na(use_type)) %>%
  pivot_wider(values_from = "use_type", names_from = "report_name") %>% 
  select(-category) %>%
  arrange(use_group) %>%
  rename(`Use group` = use_group,
         `Urban Water Management Plan` = UWMP,
         `Monthly Urban Water Conservation Report` = CR,
         `Electronic Annual Report` = EAR,
         `Water Loss Audit` = WLR)

DT::datatable(use_type_mapping_table, rownames = F) %>%
  formatStyle('Urban Water Management Plan', background = styleEqual(NA, 'lightgray')) %>%
  formatStyle('Monthly Urban Water Conservation Report', background = styleEqual(NA, 'lightgray')) %>%
  formatStyle('Electronic Annual Report', background = styleEqual(NA, 'lightgray')) %>%
  formatStyle('Water Loss Audit', background = styleEqual(NA, 'lightgray')) 
```

## Water supply groups

Table 2 below shows the mapping for water supply groups across the UWMP, CR, EAR, and WLR. 
The UWMP requires the most supply groups to be reported on. The EAR requires
some of the same groups as the UWMP but not all. The WLR and CR require reporting
of water supply at a coarse scale.

The water use and supply mapping tables will be an important guide throughout this
analysis.

```{r supply_mapping, echo = F}
supply_type_mapping_table <- filter(supply_type_lookup, !is.na(use_type), !grepl("total", use_group)) %>%
  pivot_wider(values_from = "use_type", names_from = "report_name") %>% 
  select(-category) %>%
  arrange(use_group) %>%
  rename(`Use group` = use_group,
         `Urban Water Management Plan` = UWMP,
         `Monthly Urban Water Conservation Report` = CR,
         `Electronic Annual Report` = EAR,
         `Water Loss Audit` = WLR)

DT::datatable(supply_type_mapping_table, rownames = F) %>%
  formatStyle('Urban Water Management Plan', background = styleEqual(NA, 'lightgray')) %>%
  formatStyle('Monthly Urban Water Conservation Report', background = styleEqual(NA, 'lightgray')) %>%
  formatStyle('Electronic Annual Report', background = styleEqual(NA, 'lightgray')) %>%
  formatStyle('Water Loss Audit', background = styleEqual(NA, 'lightgray')) 
```

```{r demand_data, include = F}
demand_analysis_raw <- demand_use_type %>%
  left_join(use_type_lookup) %>%
  filter(year == 2020) %>%
  mutate(year = as.factor(year),
         supplier_id = case_when(supplier_id == "CA2810003" ~ "Agency 1",
                                 supplier_id == "CA3710023" ~ "Agency 2",
                                 supplier_id == "CA3010073" ~ "Agency 3")) 
demand_analysis <- demand_analysis_raw %>%
  group_by(supplier_id, report_name, use_group, year) %>%
  summarize(volume_af = sum(volume_af))

total_demand <- filter(demand_analysis_raw, use_type != "recycled", !is.na(use_group)) %>%
  group_by(supplier_id, report_name) %>%
  summarize(total_demand = sum(volume_af))

demand_analysis_percent <- filter(demand_analysis_raw, use_type != "recycled", !is.na(use_group)) %>%
  group_by(supplier_id, report_name, use_group, year) %>%
  summarize(volume_af = sum(volume_af)) %>%
  left_join(total_demand) %>%
  mutate(percent_total = volume_af/total_demand)
  
# 10 demand use groups
use_group_list <- unique(demand_analysis$use_group)

# report_colors <- c("WLR" = "#1b9e77",
#                    "CR" = "#d95f02",
#                    "EAR" = "#7570b3",
#                    "UWMP" = "#e7298a")

report_colors <- c("WLR" = "#9A8822",
                   "CR" = "#74A089",
                   "EAR" = "#F8AFA8",
                   "UWMP" = "#FDDDA0")

use_type_stacked_bar <- function(dat) {
  ggplot(dat, aes(x = year, y = volume_af, fill = report_name)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = report_colors) +
  facet_wrap(~supplier_id, scales = "free_y") +
  labs(x = "") +
  theme_minimal() +
  theme(legend.title = element_blank())
}

use_type_percent_bar <- function(dat) {
  ggplot(dat, aes(x = year, y = percent_total, fill = report_name)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = report_colors) +
  facet_wrap(~supplier_id, scales = "free_y") +
  labs(x = "") +
  theme_minimal() +
  theme(legend.title = element_blank())
}
```

# Analysis themes {.tabset}

## Disconnect between data available (and when it is available) and reporting requirements

The CR requires suppliers to report water use and supply data monthly and the EAR
requires an annual submission of data aggregated by month. The UWMP is required 
every five years and the WLR is required annually.

Monthly reporting of water consumption data, particularly for the CR, is challenging 
because some suppliers may collect consumption data based on bimonthly meter read
cycles or may not collect data for a particular category. The quotes from two retailers
below describe the challenges in meeting monthly reporting requirements for water use data
and the different approaches that are used to meet the requirements.

"At the customer level, it's every two months; bimonthly for most customers. So you don't
have real time data for that; and so, there's a little bit of an issue when, like we
do a monthly report and they want to know residential use. We use, you know, an equation
to guess based on a ratio from last year or something to determine that; we just don't have
real time data for metered billed use. But we have it for production, so there's a little bit
of inconsistency there in terms of the time, the time periods that are available for us." 
(Source: Retailer 2022).

"What they are reporting to the state is not what is being asked for. And it is pretty common
knowledge among water agencies that's not what you're doing. We don't report water use at all.
Honestly, we report production, and we are hoping to change that, but we honestly just don't
have the staff or the capabilities to do it accurately."
(Source: Retailer 2022).

In Figure 1, percent of total water use reported by water use category is compared across
the CR, UWMP, EAR, and WLR. The CR asks for water use to be reported by each of the
categories shown (residential, agricultural and irrigation, CII, and losses) though 
that data is not always reported even when it exists for other reports. For instance,
Agency 1 reports agricultural use and losses on the EAR and UWMP but reports neither 
on the CR. While differences exist across all reports, the CR is often the outlier, 
and where the biggest differences exist.

This suggests that for some suppliers data reported on the CR represents an estimate
rather than actual volumes. This characteristic distinguishes the CR from the UWMP,
EAR, and WLR. 

The disconnect between state reporting requirements and the reality of local supplier
data collection and analysis methods extends beyond the CR. For instance, one supplier
discussed the timing expectations of the upcoming Annual Water Supply and Demand Assessment:

"Now have the Annual Water Supply and Demand Assessment Reporting requirement, right.
This was happening July 1st and and they assume that every district does it the same 
way, you know, we're all gonna know our projection end of June or all this for the next 
year and stuff. And we don't do it that way."
(Source: Retailer 2022).

```{r disconnect, echo = F}
filter(demand_analysis_percent, use_group %in% c("residential", "agricultural irrigation",
                                         "commerical industrial institutional", "losses")) %>%
  mutate(use_group = case_when(use_group == "residential" ~ "Residential",
                               use_group == "agricultural irrigation" ~ "Agricultural",
                               use_group == "commerical industrial institutional" ~ "CII",
                               use_group == "losses" ~ "Losses"),
         percent_total = percent_total * 100) %>%
  ggplot(aes(x = year, y = percent_total, fill = report_name)) +
  geom_bar(stat = "identity", position = position_dodge(width = 1, preserve = "single")) +
  scale_fill_manual(values = report_colors) +
  facet_grid(use_group~supplier_id, scales = "free_y") +
  labs(x = "",
       y = "Percent of total use reported") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "bottom",
        strip.text.y = element_text(size = 8)) 

```

## Redundancy across reports

Table 1 and 2 provide an interpretation of the alignment in data elements across
reports. While the names and definitions may vary, the data reported are very similar
indicating that for all intents and purposes they are interpreted as the same. 
Regardless of the intention of reporting requirements, this suggests that there 
is redundancy across reports. 

This redundancy may not always apparent when comparing raw volumes because there may
be differences in the exact values being reported on each report as discussed above.
When comparing use types across reports by percentage of total demand, in some cases,
the redundancy becomes more apparent.

For instance, Agency 1 reports different volumes of commercial, industrial and institutional
consumption across reports where the values reported on the EAR and UWMP are more similar than
on the CR. CII demand reported on the CR is 23% different than the UWMP and 20% different than the EAR.
However, when comparing this use type as a percentage of total demand reported, all reports are very similar 
(within 1% difference).

```{r redundancy, echo = F}
example_1_raw <- filter(demand_analysis, use_group == "commerical industrial institutional", supplier_id == "Agency 1") %>%
  ggplot(aes(x = year, y = volume_af, fill = report_name)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = report_colors) +
  labs(y = "Commerical industrial institutional volume (af)") +
  labs(x = "") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "none")

example_1_percent <- filter(demand_analysis_percent, use_group == "commerical industrial institutional", supplier_id == "Agency 1") %>%
  ggplot(aes(x = year, y = percent_total, fill = report_name)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = report_colors) +
  labs(y = "Commerical industrial institutional volume (af)") +
  labs(x = "") +
  theme_minimal() +
  theme(legend.title = element_blank())

gridExtra::grid.arrange(example_1_raw, example_1_percent, nrow = 1)

```


insert sentence about comparing percentages to actual values. values may be different
but percent of total reported are similar. include a few example.

Agency 2 - residential. values off but percentages really close
Agency 1 - cii, values off but percentages close

```{r, include = F}
example_1 <- filter(demand_analysis_percent, use_group == "residential", supplier_id == "Agency 2") 

example_1_format <- select(example_1, report_name, volume_af) %>%
  pivot_wider(names_from = "report_name", values_from = "volume_af") %>%
  mutate(type = "volume_af") %>%
  bind_rows(select(example_1, report_name, percent_total) %>%
  pivot_wider(names_from = "report_name", values_from = "percent_total") %>%
  mutate(type = "percent_total")) %>%
  mutate(percent_diff = (abs(CR-UWMP)/(0.5 * (CR+UWMP)))*100)

example_2 <- filter(demand_analysis_percent, use_group == "commerical industrial institutional", supplier_id == "Agency 1")

example_2_format <- select(example_2, report_name, volume_af) %>%
  pivot_wider(names_from = "report_name", values_from = "volume_af") %>%
  mutate(type = "volume_af") %>%
  bind_rows(select(example_2, report_name, percent_total) %>%
  pivot_wider(names_from = "report_name", values_from = "percent_total") %>%
  mutate(type = "percent_total")) %>%
  mutate(percent_diff_cr_uwmp = (abs(CR-UWMP)/(0.5 * (CR+UWMP)))*100,
         percent_diff_cr_ear = (abs(CR-EAR)/(0.5 * (CR+EAR)))*100,
         percent_diff_ear_uwmp = (abs(EAR-UWMP)/(0.5 * (EAR+UWMP)))*100)
```

## Resources

The resources available to local agencies to meet state reporting requirements varies
by agency. While there are nuances that exist within each agency, one broad categorization
is that generally smaller agencies have fewer staff, resources, and capacity compared
to larger agencies. 

At smaller agencies staff may be "wearing many hats" and involved in many aspects of
state reporting requirements where at larger agencies roles and responsibilities may
be more divided involving many people spread across different units or divisions. 
Larger agencies may have the capacity to develop processes and tools to streamline
reporting; however, there may be challenges in coordinating across divisions. Smaller
agencies may have challenges acquiring resources to develop streamlined processes.

# Analysis of water demand by use group {.tabset}

## Residential

Residential water use is reported on the UWMP and EAR through single family use
and multi-family use. On the CR residential use is reported as a percentage of total
potable water use. The WLR does not require residential water use to be reported.

```{r residential}
# Residential
# Example of data quality issue
# Moulton residential (sf) is huge for the EAR in 2020
# why EAR missing for Santa Fe in 2020
# wishlist pull more years for EAR
filter(demand_analysis, use_group == "residential") %>%
  use_type_stacked_bar() +
  labs(y = "Residential volume (af)")

filter(demand_analysis_percent, use_group == "residential") %>%
  use_type_percent_bar() +
  labs(y = "Residential volume (af)")
# TODO how best to show definitions or if they are needed here?
# filter(data_dictionary, `Definition group` == "Residential use type")
```

CR is greater than UWMP in 2 cases. Residential use is reported as a percentage
of total potable water rather than a volume. 

*Interview insight* The tight timeline to report data for the CR may result in data quality
issues. There may not be enough time for QA/QC checks and numbers may not be final.
In some cases, supply data rather than use data is reported on the CR. 

*Data quality* Agency 3 has abnormally large values for the EAR in 2020. This trend
persists to other use types. Agency 2 did not report any data for the 2020 EAR.

## Agriculture and irrigation

Agricultural and irrigation water use is reported on the UWMP, EAR, and CR.

```{r }
# Ag/irrigation
# City of Napa reports on the EAR/UWMP but not CR
filter(demand_analysis, use_group == "agricultural irrigation") %>%
  use_type_stacked_bar() +
  labs(y = "Agricultural irrigation volume (af)")

filter(demand_analysis_percent, use_group == "agricultural irrigation") %>%
  use_type_percent_bar() +
  labs(y = "Agricultural irrigation volume (af)")

# TODO how best to show definitions
# filter(data_dictionary, `Definition group` == "Agricultural irrigation use type")
```

Why doesn't Agency 1 report agricultural use on the CR when they report it on EAR and UWMP?
- In CR they note that their data is collected on a bimonthly meter read cycle
- Are they unable to distinguish this use type from others and it is lumped in with CII?

For both residential and agriculture use types, Agency 1 reports the exact same value on the UWMP and EAR.

## Commercial, industrial, institutional

Commercial, industrial, and institutional are reported separately on the UWMP, whereas
they are reported together on the CR. The EAR requires "commercial/industrial", and 
"industrial."

```{r}
# commerical industrial institutional
# Moulton Niguel is huge
filter(demand_analysis, use_group == "commerical industrial institutional") %>%
  use_type_stacked_bar() +
  labs(y = "Commerical industrial institutional volume (af)") 

filter(demand_analysis_percent, use_group == "commerical industrial institutional") %>%
  use_type_percent_bar() +
  labs(y = "Commerical industrial institutional volume (af)")

#filter(data_dictionary, `Definition group` == "CII use type")
```
Agency 2
- CII for CR is much larger than the UWMP. Is there a reason why this might be the case?

## Losses

Losses are reported for the UWMP, CR, and WLR, and are not reported on the EAR.

```{r}
# losses
# City of napa does not report on CR, santa fe UWMP dwarfs CR
filter(demand_analysis, use_group == "losses") %>%
  use_type_stacked_bar()
  labs(y = "Losses demand volume (af)")
  
filter(demand_analysis_percent, use_group == "losses") %>%
  use_type_percent_bar()
  labs(y = "Losses demand volume (af)")
```
Agency 1 does not report losses on the CR and Agency 2 reports a very low number.
Do these agencies not have the capacity to report losses on a monthly basis?

```{r, include = F}
# TODO make recycled its own category - is it supply or is it demand?
# recycled
# only data is on CR, make sure we aren't losing data
filter(demand_analysis, use_group == "recycled") %>%
  use_type_stacked_bar() +
  labs(y = "Recycled demand volume (af)")

```

## Landscape

The UWMP and EAR report landscape water use. The CR and WLR do not include landscape
as a separate use category.

```{r}
# landscape
# Moulton Niguel EAR is huge
filter(demand_analysis, use_group == "landscape") %>%
  use_type_stacked_bar() +
  labs(y = "Landscape demand volume (af)")

filter(demand_analysis_percent, use_group == "landscape") %>%
  use_type_percent_bar() +
  labs(y = "Landscape demand volume (af)")
```

## Sales, transfers, exchanges

The UWMP, EAR, and WLR require sales, transfers and exchanges to be reported.
Sales, transfers and exchanges are reported as one group for the UWMP. Water that is
sold is reported on the EAR and water that is exported is reported on the WLR.

```{r}
# sales transfers exchanges to other agencies
filter(demand_analysis, use_group == "sales transfers exchanges to other agencies") %>%
  use_type_stacked_bar() +
  labs(y = "Sales transfers exchanges to other agencies volume (af)")

filter(demand_analysis_percent, use_group == "sales transfers exchanges to other agencies") %>%
  use_type_percent_bar() +
  labs(y = "Sales transfers exchanges to other agencies volume (af)")
```

# Analysis of water supply by group {.tabset}

Groundwater and nonpotable supply categories are required buy are not included here
as the agencies in this analysis do not report any water supplied through these
categories.

```{r supply_data, include = F}
supply_analysis <- supply_use_type %>%
  left_join(supply_type_lookup) %>%
  filter(year == 2020) %>%
  mutate(year = as.factor(year),
         supplier_id = case_when(supplier_id == "CA2810003" ~ "Agency 1",
                                 supplier_id == "CA3710023" ~ "Agency 2",
                                 supplier_id == "CA3010073" ~ "Agency 3"))

total_supply <- filter(supply_analysis, use_group != "recycled water", !is.na(use_group)) %>%
  group_by(supplier_id, report_name) %>%
  summarize(total_supply = sum(volume_af))

supply_analysis_percent <- left_join(filter(supply_analysis, use_group != "recycled water", !is.na(use_group)), 
                                     total_supply) %>%
  mutate(percent_total = volume_af/total_supply)
  
# 10 demand use groups
supply_group_list <- unique(supply_analysis$use_group)
```

## Imported/purchased

```{r}
filter(supply_analysis, use_group == "imported/purchased") %>%
  use_type_stacked_bar() +
  labs(y = "Imported/purchased volume (af)")

filter(supply_analysis_percent, use_group == "imported/purchased") %>%
  use_type_percent_bar() +
  labs(y = "Imported/purchased volume (af)")
```
Across all three reports where imported/purchased water is a supply category,
Agency 3 reports 100% of water supplied through this source. The actual volume
supplied varies slightly though across reports.

Agency 1 and 2 report different primary sources across reports, shown through
the categories below.

## Surface water

```{r}
filter(supply_analysis, use_group == "surface water") %>%
  use_type_stacked_bar() +
  labs(y = "Surface water volume (af)")

filter(supply_analysis_percent, use_group == "surface water") %>%
  use_type_percent_bar() +
  labs(y = "Surface water volume (af)")
```
Both Agency 1 and 2 report 100% of water sourced from surface water on the EAR.

## Storage

```{r}
filter(supply_analysis, use_group == "storage") %>%
  use_type_stacked_bar() +
  labs(y = "Storage volume (af)")

filter(supply_analysis_percent, use_group == "storage") %>%
  use_type_percent_bar() +
  labs(y = "Storage volume (af)")
```

## Own sources

```{r}
filter(supply_analysis, use_group == "own sources") %>%
  use_type_stacked_bar() +
  labs(y = "Own sources volume (af)")

filter(supply_analysis_percent, use_group == "own sources") %>%
  use_type_percent_bar() +
  labs(y = "Own sources volume (af)")
```
Agency 1 and 2 report 100% of water sourced through own sources on the WLR.

# Issues

Characteristics of local agencies relative to state reporting requirements
Data collection
- Frequency and scale of data collected do not align with reporting requirements,
particularly the CR monthly reporting
- Examples: Bimonthly meter cycle does not align with monthly reporting requirements,
budget-based rate structures compared to use types defined by land use, some water uses
require manual reading, timeline of the Annual Water Supply and Demand Assessment
does not align with internal timelines for data reporting and planning

Data processing
- Timing of QA/QC does not align with reporting requirements, particularly the CR
monthly reporting
- Examples: Reconciling incoming/outgoing water users may not happen in time for
monthly reports, data may not be ready or available so estimates are applied instead,
even in cases where water use is automatically collected
through meter reads there may be QA/QC issues that require manual checks

Data analysis
- Different methodologies are used to summarize data across agencies
- Different methodologies and data are used within an agency across reports
- Examples: Water production data may be used for the monthly reports because
billing data are not yet available, estimates may be used in cases where data
are not collected at that scale or maybe not reported at all, within an agency
different units may be responsible for different reports and may use methods
that are different than others resulting in different numbers

Characteristics of state reporting requirements
High volume
Challenging timelines
High complexity
Inconsistency of definitions across reports
Redundant information across reports
Lack of clarity about what data are being used for

There is a disconnect between state reporting requirements and the reality of what data
are collected by suppliers and how that data is collected and reported. This results
in a lack of data quality with unknown error and uncertainty. In order to improve 
data quality and utility either local agencies need to adjust their processes to
meet reporting requirements or state reporting requirements need to be adjusted
to better meet the reality of local suppliers. There may also be opportunities
for movement from the state and local side to meet somewhere in the middle.

# Recommendations

Aligning state reporting requirements with data available locally (currently or could be developed in future), 
and local data collection methods. This may require developing groups or 
clusters of suppliers that are most similar to develop appropriate solutions.

Developing a data management system that facilitates automated and standardized
reporting of water use and supply data to meet the volume, timing, and complexity 
of state reporting requirements.
- Organized by state or third-party
- Development and building would be funded by the state
- Maintenance would be funded by local suppliers
- System would be built for state (or others) to pull in data elements and summarize as
needed for the specific report.



